---
title: "DATA 599 - Survival Analysis HW4 - Charles Hanks"
date: "Due: February 17, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(skimr)
library(survival)
library(survminer)
library(fitdistrplus)
```

DATASET: genfan (preloaded with survival package)

This dataset contains the failure time (hrs of service) of 70 diesel engine fans.

Please complete the following:

## A (20)
Fit a KM curve to the data.  Make sure your graph has labels and a title.

```{r}
# your code below

```

## B (15)
Fit an exponential distribution to the data (accounting for censoring).  What is the estimate for the rate parameter?

```{r}

```
*ANSWER:* [FILL IN]

## 3 (15)

Fit the Weibull distribution to the data (accounting for censoring).  What are the estimates for the shape and scale parameters?

```{r}

```
*ANSWER:* [FILL IN]

## D (15)
Fit the lognormal distribution to the data (accounting for censoring).  What are the estimates for the location and sd parameters?

```{r}

```
*ANSWER:* [FILL IN]

## E (20)
Create side-by-side QQ plots for the 3 fits using cdfcompcens.

```{r}

```
*ANSWER:* [FILL IN]

### F (10)
Which distribution best fits the data?  Justify your answer.

*ANSWER:* [FILL IN]

### F
Suppose the stated warranty for the fans is 1000 hours.  Using your distribution fit, what's the probability of a given diesel fan failing in warranty?

```{r}

```
*ANSWER:* [FILL IN]