---
title: "DATA 599 - Survival Analysis HW7 - Charles Hanks"
date: "Due: April 7, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/charleshanks/desktop/msds/spring_23/SA/datasets')

library(tidyverse)
library(skimr)
library(survival)
library(survminer)
library(broom)
library(autoplotly) 
```

** CHANGE YOUR NAME IN THE TITLE AND COMPLETE THE FOLLOWING: **

# DATASET: pharmacoSmoking-asaur.txt

This dataset, leveraged from *Applied Survival Analysis Using R* contains the following variables:

- id  (patient ID number) 
- ttr (Time in days until relapse)
- relapse (Indicator of return to smoking) 
- grp (Randomly assigned treatment group with levels combination or patchOnly)
- age (Age in years at time of randomization)
- gender (Female or Male) 
- race (black, hispanic, white, or other)
- employment (ft=full-time, pt=part-time, or other)
- yearsSmoking (Number of years the patient had been a smoker)
- levelSmoking (heavy or light)
- ageGroup2 (Age group with levels 21-49 or 50+)
- ageGroup4 (Age group with levels 21-34, 35-49, 50-64, or 65+)
- priorAttempts (The number of prior attempts to quit smoking)
- longestNoSmoke (The longest period of time, in days, that the patient has previously gone without smoking)

The time to relapse (ttr) is the time variable of interest, and the relapse variable is the censoring variable.  Please read in the dataset and do the following:

## A
We need to do some wrangling.  Replace all 0's in the `ttr` column with 0.00001 (or a similarly tiny #).  Also, change the order of the race variable levels to {white, black, hispanic, other}.

```{r}
ds = read_csv('pharmacoSmoking-asaur.txt')

ds = ds %>% mutate(ttr = ifelse(ttr == 0, 0.0001,ttr)) %>% 
       mutate(race = factor(race, levels = c('white', 'black','hispanic','other')))

```

## B
Create separate KM survival curves for the following relationships:
- ttr vs. grp
- ttr vs. gender
- ttr vs. race
- ttr vs. levelSmoking
- ttr vs. ageGroup4

```{r}
#testing on first variable
#km.grp = survfit(Surv(ds$ttr,ds$relapse) ~ ds$grp)
#ggsurvplot(fit = km.grp, data = ds, risk.table = F, conf.int = T)

#creating list of var names to iterate over
vars = list('grp', 'gender', 'race', 'levelSmoking', 'ageGroup4')

#iterating through list of vars to generate separate km curve for each km table
for(var in vars){
  str = paste0('survfit(Surv(ds$ttr,ds$relapse) ~ ds$', var,')')
  km.fit = eval(parse(text = str))
  print(ggsurvplot(fit = km.fit, data = ds, risk.table = F, conf.int = F) + labs(title = paste('Time to Relapse vs. ', var,sep ='')))
}
```

## C
Create a joint Weibull model with the variables in part A using `survreg`.  Name this model `weibullMod`.

```{r}
weibullMod = survreg(Surv(ttr,relapse)~ grp + gender + race + levelSmoking + ageGroup4, data = ds)
```

## D
Run `broom::tidy(weibullMod)` on the model.  Which variables are statistically significant?  (Hint: which variables have a p-value < 0.05?)

```{r}
broom::tidy(weibullMod)
```

**ANSWER:** The statistically significant variables are grp and ageGroup4. 

## E
Run a Cox proportional hazards model with the same variables in part A using `coxph`.  Name this model `coxMod`.

```{r}
coxMod=coxph(Surv(time,status)~trt + celltype + karno + diagtime + age + prior,data=veteran)
coxMod = 
```

## F
Run `summary(coxMod)`.  What is the interpretation of the grp (i.e. treatment) coefficient?  Hint: look at the coefficient in the exp(coef) column.  What does it mean to be >1?

```{r}

```

**ANSWER:** [fill in]

## G
What is the interpretation of the smoking level coefficient?

**ANSWER:** [fill in]

## H
We need to check the model diagnostics to verify that this model satisfies assumptions of the Cox PH model.  Uncomment and run the following code:

```{r}
# ggcoxdiagnostics(coxMod, type = "dfbeta", linear.predictions = FALSE, ggtheme = theme_minimal())

# ggcoxdiagnostics(coxMod, type = "deviance",  linear.predictions = FALSE, ggtheme = theme_minimal())
```

Examine the plots.  Are there any influential points?

**ANSWER:** [fill in]

## I
Run the following code to check the proportional hazards assumption:

```{r}
#coxZPH=cox.zph(coxMod)
#ggcoxzph(coxZPH,font.x=10,font.y=10,font.main=12)
```

Based on the Schoenfeld residuals, is the proportional hazards assumption satisfied?

**ANSWER:** [fill in]

## J
Using the variable names in the dataset, create a data frame for a 60yo Black male who is a light smoker.  One row should correspond to grp="patchOnly" and the second row should correspond to the combination treatment.  (Hint: expand.grid).  Name this data frame `newData`.

```{r}

```

## K
Uncomment and run the following code to create the survival tables for the cox PH curves:

```{r}
# coxSurvTables=survfit(coxMod, data = veteran, newdata = newData, conf.type = "none")
```

Then use the survfit object and `ggsurvplot` to create survival curves for a 60yo Black male.  There should be a curve for each treatment (patchOnly, combination).

```{r}

```
