---
title: "DATA 599 - Survival Analysis HW4 - Charles Hanks"
date: "Due: February 17, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(skimr)
library(survival)
library(survminer)
library(fitdistrplus)
```

DATASET: genfan (preloaded with survival package)

This dataset contains the failure time (hrs of service) of 70 diesel engine fans.

Please complete the following:

## A (20)
Fit a KM curve to the data.  Make sure your graph has labels and a title.

```{r}
km.genfan = survfit(Surv(genfan$hours,genfan$status)~1)
summary(km.genfan)
ggsurvplot(fit = km.genfan, data = genfan, palette = "forestgreen", risk.table = F, conf.int = T) + 
  labs(title = "Survival Curve for Diesel Engine Fans", 
       x = "Time (Hours)",
       y = "Survival Probability") 

```

## B (15)
Fit an exponential distribution to the data (accounting for censoring).  What is the estimate for the rate parameter?

```{r}
genfan = genfan %>% mutate(left = hours, 
                           right = ifelse(status==1, hours, NA))

interval.genfan = genfan %>% dplyr::select(left,right)

exp.fit = fitdistcens(data.frame(interval.genfan),"weibull",fix.arg=list(shape=1))
summary(exp.fit)

#rate of exponential distribution is equal to 1 / scale of weibull distribution: 
rate = 1/as.numeric(unlist(exp.fit[1]))

```
*ANSWER:* Estimate for rate parameter is 5.998467e-05

## 3 (15)

Fit the Weibull distribution to the data (accounting for censoring).  What are the estimates for the shape and scale parameters?

```{r}
weibull.fit = fitdistcens(data.frame(interval.genfan), "weibull")
summary(weibull.fit)
```
*ANSWER:* Estimate for shape parameter: 1.058, estimate for scale: 26301.309

## D (15)
Fit the lognormal distribution to the data (accounting for censoring).  What are the estimates for the location and sd parameters?

```{r}
lnorm.fit = fitdistcens(data.frame(interval.genfan), "lnorm")
summary(lnorm.fit)
```
*ANSWER:* location estimate: 10.146, sd estimate: 1.681. 

## E (20)
Create side-by-side QQ plots for the 3 fits using cdfcompcens.

```{r}
qqcompcens(list(exp.fit,weibull.fit,lnorm.fit),
            legendtext=c("Exponential","Weibull","Lognormal"),
            xlab="Failure Time (Hours)",
            plotstyle = "ggplot")
```
*ANSWER:* [FILL IN]

### F (10)
```{r}
aic = c(exp.fit$aic, weibull.fit$aic, lnorm.fit$aic)
fits = c("exponential", "weibull", "lognormal")

cbind(fits,data.frame(aic))

```

Which distribution best fits the data?  Justify your answer.

*ANSWER:* The lognormal distribution is the best fit because it has the lowest AIC. Also one can see in the QQ plot above that the lognormal data deviates the least from the 45 degree line. 

### F
Suppose the stated warranty for the fans is 1000 hours.  Using your distribution fit, what's the probability of a given diesel fan failing in warranty?

```{r}
plnorm(1000, 10.145528, 1.680757, lower.tail = TRUE)

```
*ANSWER:* The probability of a given diesel fan failing in warranty is .027. There is about a 3% chance that the diesel fan will fail within warranty. 